Time-series modeling of epidemics in complex populations: detecting changes in incidence volatility over time

### 2024-12-19 ms-3/ms.tex
Update

\section*{Methods}
By definition incidence volatility is fast relative to broadscale epidemic dynamics. 
Consequently, in order to estimate incidence volatility we first model incidence at broad spatiotemporal scales using natural splines~\citep{perperoglou2019review} fitted within a moving window
\begin{equation}
  \log\left(\dfrac{\mu_t}{N}\right)  = \sum_{j=1}^J \beta_j^{(t)} h_j(t)
\end{equation}
where $N$ represents population size, $h_j(t)$ are basis functions, $J$ is the degrees of freedom for the splines, and $\beta_j^{(t)}$ are fitted parameters for a symmetrical window of half-width $\Delta$, centered at $t$, i.e., extending from $t-\Delta$ to $t + \Delta$. 
The degrees of freedom to be used for the splines, and the width of the moving window will depend on the application.
Explaination of the specific choices we used $J$ and $\Delta$ for our application to COVID-19 cases in US counties is described below.

@@

We estimate $\theta_t$ given observed incidence using an iteratively reweighted least-squares procedure implemented via the NBPSeq R package~\citep{di2015nbpseq}. 
For each time window, a series of $\mu_s$ from $s = t-\Delta$ to $s = t + \Delta$ is estimated using the spline, and a single value of $\theta_t$ is fitted. [more detail needed]

>>>>

\section*{Methods}
By definition incidence volatility is fast relative to broadscale epidemic dynamics. 
Consequently, in order to estimate incidence volatility we first model incidence at broad spatiotemporal scales using natural splines~\citep{perperoglou2019review}. To allow for diverse shapes in the broadscale epidemic dynamics, these are fitted within a moving window
\begin{equation}
  \log\left(\dfrac{\mu_t}{N}\right)  = \sum_{j=1}^J \beta_j^{(t)} h_j(t)
\end{equation}
where $N$ represents population size, $h_j(t)$ are basis functions, $J$ is the degrees of freedom for the splines, and $\beta_j^{(t)}$ are fitted parameters for a symmetrical window of half-width $\Delta$, centered at $t$, i.e., extending from $t-\Delta$ to $t + \Delta$. 
The degrees of freedom to be used for the splines, and the width of the moving window will depend on the application.
Explaination of the specific choices we used $J$ and $\Delta$ for our application to COVID-19 cases in US counties is described below.

@@

We estimate $\theta_t$ given observed incidence using an iteratively reweighted least-squares procedure implemented via the NBPSeq R package~\citep{di2015nbpseq}. 
For each time window, a series of $\mu_s$ from $s = t-\Delta$ to $s = t + \Delta$ is estimated using the spline, and a single value of $\theta_t$ is fitted.
